Transfer of Learned Opponent Models in Zero Sum Games

### 2020-01-15 Report/draft_report_v1.Rmd
finalise v1 of report

---
title             : "Transfer of Learned Opponent Models in Zero Sum Games"
shorttitle        : "draft"

@@

author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

@@

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
linkcolor         : "blue"
mask              : no
draft             : no

class             : "man"
output            : papaja::apa6_pdf
header-includes   :
  - \usepackage[document]{ragged2e}
---

```{r load-packages, include = FALSE}
#library("papaja")
require(knitr)
require(citr)
# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(tidyr)
library(dplyr)
library(MASS)
library(ggpubr)
library(PairedData)
library(multcompView)
library(lsmeans)
```

@@

```{r texreport}
texreport <- function(object,effect.name,...) UseMethod("texreport")
texreport.default <- function(object,effect.name,...) return(NULL)

texreport.summary.merMod <- function(object,effect.name,print.estimate=TRUE) {
  if(effect.name %in% rownames(object$coefficients)) {
    tab <- object$coefficients
    estimate <- tab[effect.name,"Estimate"]
    se <- tab[effect.name,"Std. Error"]
    Tval <- tab[effect.name,"t value"]
    df <- tab[effect.name,"df"]
    pval <- tab[effect.name,"Pr(>|t|)"]
    if(pval < .001) ptext <- "$p < .001$" else {
      if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
        ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
      }
    }
    txt <- paste("$b = ",round(estimate,2),"$, $SE = ",round(se,2),"$, $t(",round(df,2),") = ",round(Tval,2),"$, ",sep="")
    txt <- paste(txt,ptext,sep="")
  } else {
    txt <- "error"
  }
  txt
}

texreport.summary.Anova.mlm <- function(object,effect.name,effect.size=TRUE,sphericity.correction=TRUE) {
  if(effect.name %in% rownames(object$sphericity.tests) && sphericity.correction && object$sphericity.test[effect.name,"p-value"] < .05) {
    epsilon <- object$pval.adjustments[effect.name,"HF eps"]
    pval <- object$pval.adjustments[effect.name,"Pr(>F[HF])"]
  } else {
    epsilon <- 1
    pval <- object$univariate.tests[effect.name,"Pr(>F)"]
  }
  df1 <- object$univariate.tests[effect.name,"num Df"]*epsilon
  df2 <- object$univariate.tests[effect.name,"den Df"]*epsilon
  Fval <- object$univariate.tests[effect.name,"F"]
  etasq <- object$univariate.tests[effect.name,"SS"]/(object$univariate.tests[effect.name,"SS"] + object$univariate.tests[effect.name,"Error SS"])
  # generate latex output
  if(pval < .001) ptext <- "$p < .001$" else {
    if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
      ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
    }
  }
  txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
  if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
  txt <- paste(txt,ptext,sep="")
  txt
}

texreport.linearHypothesis.mlm <- function(object,effect.size=TRUE) {
  if(ncol(object$SSPH) == 1 & nrow(object$SSPH) == 1 & ncol(object$SSPE) == 1 & nrow(object$SSPE) == 1) {
    df1 <- object$df
    df2 <- object$df.residual
    Fval <- (object$SSPH/df1)/(object$SSPE/df2)
    pval <- 1-pf(Fval,object$df,object$df.residual)
    etasq <- object$SSPH/(object$SSPH + object$SSPE)
    # generate latex output
    if(pval < .001) ptext <- "$p < .001$" else {
      if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
        ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
      }
    }
    txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
    if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
    txt <- paste(txt,ptext,sep="")
  } else {
    txt <- "error"
  }
  txt
}

texreport.anova.lme <- function(object,effect.name,effect.size=FALSE) {
  if(effect.name %in% rownames(object)) {
    pval <- object[effect.name,"p-value"]
    if(is.null(pval)) pval <- object[effect.name,"Pr(>F)"]
    df1 <- object[effect.name,"numDF"]
    df2 <- object[effect.name,"denDF"]
    Fval <- object[effect.name,"F-value"]
    #etasq <- object$univariate.tests[effect.name,"SS"]/(object$univariate.tests[effect.name,"SS"] + object$univariate.tests[effect.name,"Error SS"])
    # generate latex output
    if(pval < .001) ptext <- "$p < .001$" else {
      if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
        ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
      }
    }
    txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
    #if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
    txt <- paste(txt,ptext,sep="")
    txt
  }
}

texreport.anova <- function(object,effect.name,effect.size=FALSE) {
  if(effect.name %in% rownames(object)) {
    pval <- object[effect.name,"Pr(>F)"]
    df1 <- object[effect.name,"NumDF"]
    df2 <- object[effect.name,"DenDF"]
    Fval <- object[effect.name,"F.value"]
    #etasq <- object$univariate.tests[effect.name,"SS"]/(object$univariate.tests[effect.name,"SS"] + object$univariate.tests[effect.name,"Error SS"])
    # generate latex output
    if(pval < .001) ptext <- "$p < .001$" else {
      if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
        ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
      }
    }
    txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
    #if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
    txt <- paste(txt,ptext,sep="")
    txt
  }
}

@@

```{r, include = FALSE}
#Experiment 1 raw data
dat <- read.csv("../data20180719.csv")

# transform 'winner' variable in numeric score
dat$score <- recode(dat$winner, human = 1, tie = 0, ai = -1)
# create a new variable 'block' with round 1...25 = block 1 and round 26...50 as block 2
dat$block <- factor(as.numeric(cut(dat$round,2)),labels =c("first half", "second half"))

# create a new variable "game.f" as a factor variable of games
dat$game.f <- factor(dat$game, labels = c("RPS","FWG","NUM"),levels=c("rps","fwg","numbers"))

#Group data by human_id and calculate mean score per block of each game.
dat2 <- dat %>% 
  group_by(human_id,condition,game,block,game.f) %>% 
      summarize(block_score = mean(score))

# Group data by game and ID
dat3 <- group_by(dat2, human_id,game.f,game) %>% summarise(game_score = mean(block_score))

```{r, include = TRUE}
# AFEX PACKAGE FOR RUNNING RM ANOVA AND POST HOC TESTS ############################
citation(package = "afex", lib.loc = NULL, auto = NULL)
library(afex)
a1 <- aov_car( block_score ~ game.f*block*condition + Error(human_id/(game.f*block)), dat2)
summary(a1)
(a1_nice <- nice(a1, es = attr(a1$anova_table, "es"),
  observed = attr(a1$anova_table, "observed"),
  correction = attr(a1$anova_table, "correction"), MSE = NULL,
  intercept = NULL, p_adjust_method = attr(a1$anova_table,
  "p_adjust_method"), sig_symbols = attr(a1$anova_table, "sig_symbols")))

write.table(a1_nice,file = "a1.txt", sep = ",", quote = FALSE, row.names = F)
```

```{r}
#Pair waise comparison and post hoc tests for the interaction: Game * Block using lsmeans

# game score differences statistically significant?
(ls0 <- lsmeans(a1, "game.f"))
(lsm0 <- update(pairs(ls0, reverse = TRUE), by=NULL, adjust = "holm"))

# Main effect of block 
(ls01 <- lsmeans(a1, "block"))
(lsm01 <- update(pairs(ls01, reverse = TRUE), by=NULL, adjust = "holm"))
lsmip(a1, ~block)

# Main effect of condition
(ls02 <- lsmeans(a1, "condition"))
(lsm02 <- update(pairs(ls02, reverse = TRUE), by=NULL, adjust = "holm"))

```{r}
# Learning transfer using first half scores: Pairwise comparison of first halves by game ( first half RPS vs first half WFG scores....)
(ls1 <- lsmeans(a1, "game.f", by="block"))
(lsm1 <- update(pairs(ls1, reverse = TRUE), by=NULL, adjust = "holm"))
lsmip(a1, game.f ~ block)
```

```{r}
(means.int <- lsmeans(a1, specs = c("game.f","block","condition")))

# Transfer level1: compare H1FWG with H1RPS 
con1  <- contrast(means.int, list(H1_G2vsH1_G1forlvl1 = c(-1,1,0,0,0,0,0,0,0,0,0,0), H1_G3vsH1_G1forlvl1=c(-1,0,1,0,0,0,0,0,0,0,0,0)))
summary(con1, adjust = "holm")

# Transfer level2: compare H1FWG with H1RPS 
con2  <- contrast(means.int, list(H1_G2vsH1_G1forlvl2 = c(0,0,0,0,0,0,-1,1,0,0,0,0), H1_G3vsH1_G1forlvl2=c(0,0,0,0,0,0,-1,0,1,0,0,0)))
summary(con2, adjust = "holm")
```

```{r}

dat_26 <- subset(dat,round >1 & round <7, drop =TRUE)
dat2_6 <- dat_26 %>% 
  group_by(human_id,condition,game.f,confidence,difficulty) %>% 
      summarise(early_score = mean(score))

# Check group means and SDs
group_by(dat2_6, game.f) %>%
  summarise(
    count = n(),
    mean = mean(early_score, na.rm = TRUE),
    sd = sd(early_score, na.rm = TRUE)
    )
group_by(dat2_6, game.f,condition) %>%
  summarise(
    count = n(),
    mean = mean(early_score, na.rm = TRUE),
    sd = sd(early_score, na.rm = TRUE)
    )
# plot scores per game 
ggerrorplot(dat2_6, x = "game.f", y = "early_score", group = 1, color="game.f", desc_stat = "mean_ci",palette = c("#00AFBB", "#E7B800", "#FC4E07"), order = c("RPS", "FWG","NUM"), ylab = "Score", xlab = "Games") 
```

@@

In order to explore learning transfer in strategic settings, it is generally useful to study simple games as a model of more complex interactions. More specifically, when studying learning, we need a framework that allows the study of whether and how a player takes into consideration, over time, the impact of its current and future actions on the future actions of the opponent and the future cumulative rewards. Repeated games, in which players interact repeatedly with the same opponent and have the ability to learn about the opponent's strategies and preferences (Mertens, 1990) are particularly adapted to modelling learning about the opponent.

Early literature on learning transfer in games has mostly focused on measuring the proportion of people who play normatively optimal (Nash Equilibria) or salient actions (e.g Risk Dominance) in later games, having had experience with a similar game environment previously. For instance, Ho et al. (1998) measure transfer as the proportion of players who choose the Nash Equilibrium in later p-beauty contest games, after traingn on similar games. They find there is no evidence of immediate transfer (Nash equilibrium play in the first round of the new game) but positive structural learning transfer as shown by the faster convergence to equilibrium play by experienced vs non experienced players. Camerer & Knez (2000) test learning transfer in players exposed to two games with multiple equilibria sequentially and explore the ability of players to conordinate their actions to choose a particular equilibrium in subsequent games having reached it in prior ones. They distinguish between games that are similar in a purely descriptive way, meaning similar choice labels, identity of players, format and number of action choices; and games that are similar in a strategic sense, meaning similar payoffs from combination of actions, identical equilibrium properties or significant social characteristics of payoffs such as possibility of punishment, need for fairness and cooperative vs competitive settings. They find that transfer of learning (successful coordination) occurs more readily in the presence of both descriptive and strategic similarity. If the games were only strategically similar, then the transfer was much weaker.

@@

One of the major limitations in studies of how humans adapt to computerised opponents is that they have mostly looked at the ability of players to detect and exploit action-based learning rules. They either used deterministic strategies that did not consider the human participant's play, as in the case of pre-determined mixed strategy algorithms, or strategies that only depended on the player's patterns of past plays. These strategies did not necessarily take into account the computer's own past plays, or the degree of strategic reasoning of the human player about the computer. Their play was not "human-like" in the sense that humans are not very good at playing specific mixed strategies with any precision of at detecting patterns from long sequences of past play due to cognitive constraints. It is therefore important to have agents that "play like humans", and one way of achieving that is to embed theses agents with human-like theory of mind abilities based on limited steps of recursive reasoning. Simon (1972) explains that humans have limited cognitive capacities and as such cannot be expected to solve computationally intractable problems such as finding Nash equilibria. Instead, they will try to 'satisfice' by choosing a strategy that is adequate in a simplified model of the environment, rather than an optimal one. This concept finds its natural application in 'level-k' theory, first adopted by Stahl & Wilson (1995). It posits that deviations from Nash equilibrium solutions in simple games are explained by the fact that humans have a heterogenous degree of strategic sophistication. At the bottom of the ladder, level-0 players are non-strategic and play either randomly or use a salient strategy in the game environment (Arad & Rubinstein, 2012). Level-1 players are next up the ladder of strategic sophistication and will assume all their opponents belong to the level-0 category and as such will best respond to them given this assumption. Likewise, a level-2 player will choose actions that are the best response given the belief that all opponents are exactly one level below and so on.

@@

To address some of these limitations in the existing literature on learning transfer outlined above, we propose to explore opponent modelling and its transfer with the use of computer agents possessing human-like theory of mind abilities with limited degrees of iterated reasoning. The agents will have a fixed strategy played stochastically and will either be level-1 or level-2 behavioral rule, mimicking human theory of mind abilities and the limited recursion depth they exhibit (Goodie et al., 2012). We also use a set of purely competitive zero-sum games that incentivise reasoning about the opponent without introducing social confounding factors. We measure transfer of learning about the opponent strategy between games that are similar and dissimilar. The first two games we use are identical except for action labels. In one experiment, the third game is strategically similar to the first two but descriptively different, while in a second experiment, we introduce a third game that is dissimilar to the first two in terms of payoff matrix and strategic structure while retaining, like all other games, a unique mixed strategy Nash equilibrium of random action with 0 expected reward against a Nash player.

@@

Participants were informed they would play three different games against the same computer opponent. Each participant plays all three games consecutively and in the same order described above. Participants were told that the opponent cannot cheat and will choose its actions simultaneously without knowledge of the participant's choice. A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against the opponents, players were paid a bonus at the end of the experiment that was proportional to their final score. The overall score of the players was translated into the bonus by making each point worth $£$0.02. This bonus is significant as players could increase the total payoff from the experiment by up to 60% assuming they'd won all rounds against the computer opponent.

@@

```{r feedback_rps, fig.cap = "Screenshot of the feedback at the end of a round of Rock-Paper-Scissors", fig.align = "center"}

# knitr::include_graphics("../Report/images/feedback_rps.png", dpi = 108)

@@

In order to encourage participants to think about their next choice, a countdown timer of 3 seconds was introduced at the beginning of each round. During those 3 seconds, participants could not choose an option and had to wait for the timer to run out. A small delay that changed randomly (between 0.5 and 3 seconds) was also introduced in the time it took the AI agent to give back their response to make it seem like it is also thinking about its last actions, as a way of simulating a real human opponent. After each round, the participants were given detailed feedback about their opponent actions as well as whether they won or lost the last round. Further information about the outcome of previous rounds was also visible on the game screen below the feedback area, throughout each stage game and opponent could go back many rounds to study the history of interaction. The number of wins, losses and ties were clearly shown at the top of the screen for each game, and this scoreboard was reinitialised to zero at the onset of a new stage game. An example of the interface for the rock-paper-scissors game is provided in Figure XX.

<!-- Figure\ \@ref(fig:feedback_rps) -->

@@

As a measure for learning transfer, we focus on participants’ scores in rounds 2-6, excluding the very first round for which the agent is programmed to play randomly as it has no data on prior rounds on which to build its response. A group of players with no experience of the game are expected to have scores not significantly different from 0. Any significantly positive group average scores would therefore reflect prior learning from past experiences. In figure XXX, the average score across participants by game for rounds 2-6 are plotted. Scores are also averaged across levels of condition. We test the average scores for each game against a hypothesised value of 0 for a non-experienced player using parametric one sample t-tests. As expected for the initial RPS game, the average score is not significantly different from 0 ( t(148.85) = 1.04 , p = 0.89). In FWG, the score is significantly higher than 0 ( t(148.85) = 4.58 , p < 0.0001). Unexpectedly, this is also the case for NUMBERS ( t(148.85) = 3.00, p = 0.0092).

### Moderation of learning transfer by type of opponent and game similarity

Figure XXXXX shows the marginal mean scores for each game for both level-1 and level-2 facing players. Graphically we can see that the pattern is dissimilar between level-1 and level-2 players, and we suspect transfer to be positively associated with similarity and negatively with degree of sophistication of the agent. To test these hypotheses, we run statistical tests on scores by game and opponent against the null hypothesis of 0 (no transfer). Table XXXXXXX shows the scores on rounds 2-6 for both type of opponents in each game. We also report the t-stats and p-values to test the null hypothesis that scores on these early rounds are equal to 0.  Rejecting the null hypothesis means that prior learning has been used to beat the opponent in these early rounds.  For level-1 facing players, there is evidence of learning transfer to both FWG and NUMBERS. For level-2 facing players, there is evidence for transfer to the similar game FWG (albeit scores are lower than for level-1 player) but not to the dissimilar game of NUMBERS.

TABLE OF SCORES BY OPPONENT AND GAME (PG 40)

Our results when averaging across conditions (previous section) showed that there was indeed evidence for transfer to the more dissimilar game (NUMBERS). We can see from TABLE XXXXX that this transfer is exclusively driven by level-1 facing players, as scores of level-2 facing players are close to nil. Therefore, both groups of condition can generalise to the similar game, but only those facing the less sophisticated opponent are able to generalise to the less similar game.

>>>>

---
title             : "Transfer of Learned Opponent Models in Zero Sum Games"

shorttitle        : "Draft-1"

@@

author_note: |
  <!-- Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line. -->

@@

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
linkcolor         : "blue"
mask              : no
draft             : no
fig_caption       : yes

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes   :
  - \usepackage[document]{ragged2e}
---

```{r load-packages, include = FALSE}
library(papaja)
require(knitr)
require(citr)
require(bookdown)
# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(tidyr)
library(dplyr)
library(MASS)
library(ggpubr)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
```

@@

```{r texreport}
# texreport <- function(object,effect.name,...) UseMethod("texreport")
# texreport.default <- function(object,effect.name,...) return(NULL)
# 
# texreport.summary.merMod <- function(object,effect.name,print.estimate=TRUE) {
#   if(effect.name %in% rownames(object$coefficients)) {
#     tab <- object$coefficients
#     estimate <- tab[effect.name,"Estimate"]
#     se <- tab[effect.name,"Std. Error"]
#     Tval <- tab[effect.name,"t value"]
#     df <- tab[effect.name,"df"]
#     pval <- tab[effect.name,"Pr(>|t|)"]
#     if(pval < .001) ptext <- "$p < .001$" else {
#       if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
#         ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
#       }
#     }
#     txt <- paste("$b = ",round(estimate,2),"$, $SE = ",round(se,2),"$, $t(",round(df,2),") = ",round(Tval,2),"$, ",sep="")
#     txt <- paste(txt,ptext,sep="")
#   } else {
#     txt <- "error"
#   }
#   txt
# }

# texreport.summary.Anova.mlm <- function(object,effect.name,effect.size=TRUE,sphericity.correction=TRUE) {
#   if(effect.name %in% rownames(object$sphericity.tests) && sphericity.correction && object$sphericity.test[effect.name,"p-value"] < .05) {
#     epsilon <- object$pval.adjustments[effect.name,"HF eps"]
#     pval <- object$pval.adjustments[effect.name,"Pr(>F[HF])"]
#   } else {
#     epsilon <- 1
#     pval <- object$univariate.tests[effect.name,"Pr(>F)"]
#   }
#   df1 <- object$univariate.tests[effect.name,"num Df"]*epsilon
#   df2 <- object$univariate.tests[effect.name,"den Df"]*epsilon
#   Fval <- object$univariate.tests[effect.name,"F"]
#   etasq <- object$univariate.tests[effect.name,"SS"]/(object$univariate.tests[effect.name,"SS"] + object$univariate.tests[effect.name,"Error SS"])
#   # generate latex output
#   if(pval < .001) ptext <- "$p < .001$" else {
#     if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
#       ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
#     }
#   }
#   txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
#   if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
#   txt <- paste(txt,ptext,sep="")
#   txt
# }
# 
# texreport.linearHypothesis.mlm <- function(object,effect.size=TRUE) {
#   if(ncol(object$SSPH) == 1 & nrow(object$SSPH) == 1 & ncol(object$SSPE) == 1 & nrow(object$SSPE) == 1) {
#     df1 <- object$df
#     df2 <- object$df.residual
#     Fval <- (object$SSPH/df1)/(object$SSPE/df2)
#     pval <- 1-pf(Fval,object$df,object$df.residual)
#     etasq <- object$SSPH/(object$SSPH + object$SSPE)
#     # generate latex output
#     if(pval < .001) ptext <- "$p < .001$" else {
#       if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
#         ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
#       }
#     }
#     txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
#     if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
#     txt <- paste(txt,ptext,sep="")
#   } else {
#     txt <- "error"
#   }
#   txt
# }
# 
# texreport.anova.lme <- function(object,effect.name,effect.size=FALSE) {
#   if(effect.name %in% rownames(object)) {
#     pval <- object[effect.name,"p-value"]
#     if(is.null(pval)) pval <- object[effect.name,"Pr(>F)"]
#     df1 <- object[effect.name,"numDF"]
#     df2 <- object[effect.name,"denDF"]
#     Fval <- object[effect.name,"F-value"]
#     #etasq <- object$univariate.tests[effect.name,"SS"]/(object$univariate.tests[effect.name,"SS"] + object$univariate.tests[effect.name,"Error SS"])
#     # generate latex output
#     if(pval < .001) ptext <- "$p < .001$" else {
#       if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
#         ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
#       }
#     }
#     txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
#     #if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
#     txt <- paste(txt,ptext,sep="")
#     txt
#   }
# }
# 
# texreport.anova <- function(object,effect.name,effect.size=FALSE) {
#   if(effect.name %in% rownames(object)) {
#     pval <- object[effect.name,"Pr(>F)"]
#     df1 <- object[effect.name,"NumDF"]
#     df2 <- object[effect.name,"DenDF"]
#     Fval <- object[effect.name,"F.value"]
#     #etasq <- object$univariate.tests[effect.name,"SS"]/(object$univariate.tests[effect.name,"SS"] + object$univariate.tests[effect.name,"Error SS"])
#     # generate latex output
#     if(pval < .001) ptext <- "$p < .001$" else {
#       if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
#         ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
#       }
#     }
#     txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
#     #if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
#     txt <- paste(txt,ptext,sep="")
#     txt
#   }
# }

@@

@@

In order to explore learning transfer in strategic settings, it is generally useful to study simple games as a model of more complex interactions. More specifically, we need a framework that allows the study of whether and how a player takes into consideration, over time, the impact of its current and future actions on the future actions of the opponent and the future cumulative rewards. Repeated games, in which players interact repeatedly with the same opponent and have the ability to learn about the opponent's strategies and preferences (Mertens, 1990) are particularly adapted to modelling learning about the opponent.

Early literature on learning transfer in games has mostly focused on measuring the proportion of people who play normatively optimal (Nash Equilibria) or salient actions (e.g Risk Dominance) in later games, having had experience with a similar game environment previously. For instance, Ho et al. (1998) measure transfer as the proportion of players who choose the Nash Equilibrium in later p-beauty contest games, after training on similar games. They find there is no evidence of immediate transfer (Nash equilibrium play in the first round of the new game) but positive structural learning transfer as shown by the faster convergence to equilibrium play by experienced vs non experienced players. Camerer & Knez (2000) test learning transfer in players exposed to two games with multiple equilibria sequentially and explore the ability of players to coordinate their actions to choose a particular equilibrium in subsequent games having reached it in prior ones. They distinguish between games that are similar in a purely descriptive way, meaning similar choice labels, identity of players, format and number of action choices; and games that are similar in a strategic sense, meaning similar payoffs from combination of actions, identical equilibrium properties or significant social characteristics of payoffs such as possibility of punishment, need for fairness and cooperative vs competitive settings. They find that transfer of learning (successful coordination) occurs more readily in the presence of both descriptive and strategic similarity. If the games were only strategically similar, then the transfer was much weaker.

@@

Still, one of the limitations in studies of how humans adapt to computerised opponents is that they have mostly looked at the ability of players to detect and exploit action-based learning rules. They either used deterministic strategies that did not consider the human participant's play, as in the case of pre-determined mixed strategy algorithms, or strategies that only depended on the player's patterns of past plays. These strategies did not necessarily take into account the computer's own past plays, or the degree of strategic reasoning of the human player about the computer. Their play was not "human-like" in the sense that humans are not very good at playing specific mixed strategies with any precision of at detecting patterns from long sequences of past play due to cognitive constraints. It is therefore important to have agents that "play like humans", and one way of achieving that is to embed theses agents with human-like theory of mind abilities based on limited steps of recursive reasoning. Simon (1972) explains that humans have limited cognitive capacities and as such cannot be expected to solve computationally intractable problems such as finding Nash equilibria. Instead, they will try to 'satisfice' by choosing a strategy that is adequate in a simplified model of the environment, rather than an optimal one. This concept finds its natural application in 'level-k' theory, first adopted by Stahl & Wilson (1995). It posits that deviations from Nash equilibrium solutions in simple games are explained by the fact that humans have a heterogenous degree of strategic sophistication. At the bottom of the ladder, level-0 players are non-strategic and play either randomly or use a salient strategy in the game environment (Arad & Rubinstein, 2012). Level-1 players are next up the ladder of strategic sophistication and will assume all their opponents belong to the level-0 category and as such will best respond to them given this assumption. Likewise, a level-2 player will choose actions that are the best response given the belief that all opponents are exactly one level below and so on.

@@

To address some of these limitations in the existing literature on learning transfer outlined above, we propose to explore opponent modelling and its transfer with the use of computer agents possessing human-like theory of mind abilities with limited degrees of iterated reasoning. The agents will have a fixed strategy played stochastically and will either be level-1 or level-2 behavioral rule, mimicking human theory of mind abilities and the limited recursion depth they exhibit (Goodie et al., 2012). We also use a set of purely competitive zero-sum games that incentivise reasoning about the opponent without introducing social confounding factors. We measure transfer of learning about the opponent strategy between games with varying degrees of similarity. The first two games we use are identical except for action labels. In one experiment, the third game is strategically similar to the first two but descriptively different, while in a second experiment, we introduce a third game that is dissimilar to the first two in terms of payoff matrix and strategic structure while retaining, like all other games, a unique mixed strategy Nash equilibrium of random action with 0 expected reward against a Nash player.

@@

Participants were informed they would play three different games against the same computer opponent. Each participant plays all three games consecutively and in the same order described above. Participants were told that the opponent cannot cheat and will choose its actions simultaneously without knowledge of the participant's choice. A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against the opponents, players were paid a bonus at the end of the experiment that was proportional to their final score. The overall score of the players was translated into the bonus by making each point worth £0.02. This bonus is significant as players could increase the total payoff from the experiment by up to 60% assuming they'd won all rounds against the computer opponent.

@@

In order to encourage participants to think about their next choice, a countdown timer of 3 seconds was introduced at the beginning of each round. During those 3 seconds, participants could not choose an option and had to wait for the timer to run out. A small delay that changed randomly (between 0.5 and 3 seconds) was also introduced in the time it took the AI agent to give back their response to make it seem like it is also thinking about its last actions, as a way of simulating a real human opponent. After each round, the participants were given detailed feedback about their opponent actions as well as whether they won or lost the last round. Further information about the outcome of previous rounds was also visible on the game screen below the feedback area, throughout each stage game and opponent could go back many rounds to study the history of interaction. The number of wins, losses and ties were clearly shown at the top of the screen for each game, and this scoreboard was reinitialised to zero at the onset of a new stage game. An example of the interface for the rock-paper-scissors game is provided in Figure\ \@ref(fig:feedback-rps).

```{r feedback-rps, fig.cap = "Screenshot of the feedback at the end of a round of Rock-Paper-Scissors", fig.align='center'}

knitr::include_graphics("../Report/images/feedback_rps.png", dpi = 108)

@@

@@

As a measure for learning transfer, we focus on participants’ scores in rounds 2-6, excluding the very first round for which the agent is programmed to play randomly as it has no data on prior rounds on which to build its response. A group of players with no experience of the game are expected to have scores not significantly different from 0. Any significantly positive group average scores would therefore reflect prior learning from past experiences. In Figure\ \@ref(fig: exp1-early-score), the average score across participants by game for rounds 2-6 are plotted. Scores are also averaged across levels of condition.

```{r exp1-early-score, fig.cap = "Average scores for rounds 2-6 by game in first experiment", fig.align='center'}

knitr::include_graphics("../Report/images/exp1_early_scor_game.png", dpi = 108)

We test the average scores for each game against a hypothesised value of 0 for a non-experienced player using parametric one sample t-tests. As expected for the initial RPS game, the average score is not significantly different from 0 as this is the first game and no learning is possible (t(148.85) = 1.04 , p = 0.89). In FWG, the score is significantly higher than 0 ( t(148.85) = 4.58 , p < 0.0001). This is also the case for the more dissimilar game : NUMBERS ( t(148.85) = 3.00, p = 0.0092).

Next, we explore whether learning transfer is moderated by the type of opponent and game similarity. Figure\ \@ref(fig: exp1-score-by-opp) shows the marginal mean scores for each game for both level-1 and level-2 facing players. Graphically we can see that the pattern is dissimilar between level-1 and level-2 players, and we suspect transfer to be positively associated with similarity and negatively with degree of sophistication of the agent. To test these hypotheses, we run statistical tests on scores by game and opponent against the null hypothesis of 0 (no transfer).

```{r exp1-score-by-opp, fig.cap = "Marginal means of players scores by game and type of opponents", fig.align='center'}

knitr::include_graphics("../Report/images/exp1_earlyScore_by_opp.png", dpi = 108)

Table\ \@ref(tab:table-early-scores) shows the scores on rounds 2-6 for both type of opponents in each game. We also report the t-stats and p-values to test the null hypothesis that scores on these early rounds are equal to 0.  Rejecting the null hypothesis means that prior learning has been used to beat the opponent in these early rounds.  For level-1 facing players, there is evidence of learning transfer from RPS to both FWG and NUMBERS. For level-2 facing players, there is evidence for transfer from RPS to the similar game FWG (albeit scores are lower than for level-1 player) but not to the dissimilar game of NUMBERS.

```{r ANOVA early exp1, include = FALSE, warning=FALSE}
dat2_6 <- read.csv(file = "../exp1_early_rounds.csv")

exp1_early_anova <- afex :: aov_car(
  early_score ~ game.f*condition + Error(human_id/(game.f))
  , data=dat2_6
  , type = 3)

apa_lm1 <- apa_print(exp1_early_anova)

means.int2 <- lsmeans(exp1_early_anova, specs = c("game.f","condition"))
trans26 <- summary(means.int2, infer = c(TRUE,TRUE),level = .95, adjust = "holm",ref=c("FWG","NUM"))

```
```{r table-early-scores, results = "asis"}
apa_table(trans26, caption = "Average early round scores by game and condition in first experiment")
```
